---
title: "Bove Meta Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r start, message=FALSE}
setwd("/Users/Colleen/Dropbox/Boston_Experiment/Manuscript3_metaanalysis/Code")
df <- read.csv("effectsize_26Nov18.csv") 
df$cite <- paste(df$author, df$year, sep = " ")

library("metafor")
library("MAd")
library("ggplot2")
library("tidyr")
library("plyr")
library("lme4")
theme_set(theme_bw())
dodge<-position_dodge(0.2)
```

# Create the Effect Size and Variance Data Frame
### Set up the dataframe of effect size and variance for acidification and warming ONLY studies

Effect size (yi) and variance (vi). Variables with a 1 are control, 3 represent acidification, 4 are warming, and 6 is the interaction studies. 'm' is the mean calcification rate, 'n' is sample size, and 's' is the standard error of mean calcification rate per study. 

```{r ES1, echo=TRUE, warning=FALSE, message=FALSE, tidy=TRUE}
# calculate ES and variance for ocean acidification
ES.oa<-escalc(measure="ROM", m1i=cm1, m2i=am3, sd1i=s1, sd2i=s3, n1i=n1, n2i=n3,  method="REML", data=df)

# calculate ES and variance ES for ocean warming
ES.ow<-escalc(measure="ROM", m1i=cm1, m2i=wm4, sd1i=s1, sd2i=s4, n1i=n1, n2i=n4,  method="REML", data=df)

#create DF of calculated variables of OA
acid<-data.frame("cm1"=ES.oa$cm1,"yi.a"=ES.oa$yi,"vi.a"=ES.oa$vi)

# merge together warming and acid DF 
ES.full<-merge(ES.ow,acid,by="cm1")
ES.full<-rename(ES.full,c("yi"="yi.w","vi"="vi.w"))
ES.full<-ES.full[order(ES.full$cite), ]
ES.full$yi.w<-(ES.full$yi.w)*-1
ES.full$yi.a<-(ES.full$yi.a)*-1
```

Calculate the standard deviation of the calcification mean (sd) and then the pooled standard deviation (S).

```{r sd}
# calculate sd for all 
ES.full$sd1<-ES.full$s1*sqrt(ES.full$n1)
ES.full$sd3<-ES.full$s3*sqrt(ES.full$n3)
ES.full$sd4<-ES.full$s4*sqrt(ES.full$n4)
ES.full$sd6<-ES.full$s6*sqrt(ES.full$n6)

# calculate S for all treatments
ES.full$Sa<-sqrt((ES.full$sd1^2/ES.full$n1)+(ES.full$sd3^2/ES.full$n3))
ES.full$Sw<-sqrt((ES.full$sd1^2/ES.full$n1)+(ES.full$sd4^2/ES.full$n4))
ES.full$Si<-sqrt((ES.full$sd1^2/ES.full$n1)+(ES.full$sd6^2/ES.full$n6))
```

### Calculate ES and variance on the interaction studies
These equations were taken from Harvey et al. 2013 (https://dx.doi.org/10.1002%2Fece3.516) 

```{r ESinter}
# calculate ES using Harvey et al equations for acidification, warming, and interaction studies
ES.full$lnRRa<-((ES.full$am3-ES.full$cm1)/ES.full$Sa)
ES.full$lnRRt<-((ES.full$wm4-ES.full$cm1)/ES.full$Sw)
ES.full$yi.i<-((ES.full$m6-ES.full$wm4)-(ES.full$am3-ES.full$cm1))/(2*ES.full$Si)

# calculate V for interaction studies ONLY
ES.full$vi.i<-(1/ES.full$n1)+(1/ES.full$n3)+(1/ES.full$n4)+(1/ES.full$n6)+((ES.full$yi.i^2)/(2*(ES.full$n1+ES.full$n3+ES.full$n4+ES.full$n6)))
```

Finally, write the new data frame as a new CSV file for faster use.
```{r writeCSV}
write.csv(ES.full,file="ES.full_26Nov18.csv")
```

### Create long version of ES dataframe
I am using the metafor calculated ES and variance for acidification and wamring alone, but the Harvey et al calculated ES and variance for the interactions... I am not sure if this is the best method? 
```{r start2}
# read in the saved CSV and then removed any unneeded columns
df<-read.csv("ES.full_26Nov18.csv")
df<-df[,-c(12:22)]
df<-df[,-c(17:23)]
df<-df[,-c(17:18)]
df<-df[,-2]
```

```{r datamanip}
#subset df for variance (v) values only as new df
df2b<-subset(df, select=c(X,vi.i,vi.a,vi.w))

#subset df to remove only variance (v) values as another new df
df.m <-subset(df, select=-c(vi.i,vi.a,vi.w))

# this creates df with treat and V as column headers in long format
df2b<-gather(df2b, treat, v,vi.i:vi.w)

#substitute y for v for merge later
df2b$treat<-gsub('v','y', df2b$treat)

#creates another long format df with treat and Y as headers (of Y values)
df.m <-gather(df.m, treat, y,yi.w:yi.i)
df.m $treat<-as.factor(df.m $treat)

#merge two df into one
df2<-merge(df.m, df2b, by=c("X", "treat"))

#remove any NAs
df2<-na.omit(df2)
df2$study <- paste(df2$author, df2$year, sep = "_")
df2$X<-NULL
head(df2)
```

#Fitting models to each treatment

```{r rma}
# subset data by treatment
OW<-subset(df2, treat=="yi.w")
OA<-subset(df2, treat=="yi.a")
IN<-subset(df2, treat=="yi.i")

# fit model to each treatment
ow<-rma(y,v,data=OW)
oa<-rma(y,v,data=OA)
both<-rma(y,v,data=IN)
```

###View OA model output with forest and funnel plots
```{r oa, echo=FALSE}
oa
forest(oa)
funnel(oa)
```

###View OW model output with forest and funnel plots
```{r ow, echo=FALSE}
ow
forest(ow)
funnel(ow)
```

###View interaction model output with forest and funnel plots
```{r both, echo=FALSE}
both
forest(both)
funnel(both)
```

#Fitting models by region (Florida and Belize only)
Remove regions so that only Florida and Belize studies remain
```{r region1}
df2a<-subset(df2, region!="Curacao")
df2a<-subset(df2a, region!="Little Cayman")
df2a<-subset(df2a, region!="Mexico")

# OA subset
OA2<-subset(df2a, treat=="yi.a")

# OW subset
OW2<-subset(df2a, treat=="yi.w")

# Interaction subset
IN2<-subset(df2a, treat=="yi.i")

```

Fit models to each treatment by region
```{r regionRMA, warning=FALSE}
# acidification by warming
oa.r<-rma(y,v,mods=~region, data=OA2)

# warming by regions
ow.r<-rma(y,v,mods=~region, data=OW2)

# interactions by region
in.r<-rma(y,v,mods=~region, data=IN2)
```

### View region-specific model output per treatment
**Ocean acidification**
```{r OAregion, echo=FALSE}
oa.r
```

**Ocean warming**
```{r OWregion, echo=FALSE}
ow.r
```

**Interaction of acidification and warming**
```{r intregion, echo=FALSE}
in.r
```

#Fitting mixed effects models and bootstrapping confidence intervals
Modify dataframe to remove and missing values and set seed/interations
```{r fullModel}
completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

df2 <- completeFun(df2, "y")
head(df2)

bootnum = 90 # set the number of iterations (will use 9000 for the final run)
seed = 4 #seed to make results replicatable
```

### This is the global model by treatment only
```{r}
#yourmodel <- lmer(y ~ treat + (1 | species) + (1 | author), data = df2) # this model has a singular fit
global.mod <- lmer(y ~ treat + (1 | study) , data = df2) # this model runs fine at this point, but will produce singular fits when bootstrapped

summary(global.mod) # view summary output of model
AIC(global.mod) # check the AIC

```

**Bootstrap the model and estimate the 95% confidence intervals and mean**
```{r}
out <- simulate(global.mod,nsim=bootnum,seed=seed,re.form=NULL)
boots <- apply(out,2,function(x) predict(lmer(x ~ treat + (1 | study), data = df2),re.form=NA))

boots <- cbind(predict(global.mod,re.form=NA), boots)
#estimate the mean and 95%confidence intervals for the prediction
df.a <- (cbind(df2, as.data.frame(t(apply(boots, 1, function(x) c(mean(x),quantile(x, c(0.025, 0.975))))))))

colnames(df.a)[15:17] <- c("mean", "lowerci", "upperci")
head(df.a)

```

###Plot estimated mean and CI
```{r echo=FALSE}
df.a$treat <- factor(df.a$treat, levels = c("yi.a","yi.w","yi.i"))

ggplot(df.a, aes(x=treat, y=mean))+
  theme(panel.grid.major=element_blank(), legend.position="bottom", panel.grid.minor=element_blank(), panel.background=element_blank())+
  theme(axis.line=element_line(color="black"))+
  theme(legend.key=element_blank())+
  ylab("Mean effect size (LnRR)")+
  xlab("Treatment")+
  scale_x_discrete(labels=c("yi.a" = "acidification", "yi.w" = "warming", "yi.i" = "interaction"))+
  geom_hline(yintercept=0, linetype=2, colour="black")+
  geom_errorbar(aes(ymin=lowerci, ymax=upperci, width=0.1), size=0.6)+
  geom_point(size=6, shape=16, colour="black")+
  ylim(-1.6,0.6)
```

### This is the model by region (Florida and Belize only) for all treatments
```{r}
df2a<-subset(df2, region!="Curacao")
df2a<-subset(df2a, region!="Little Cayman")
df2a<-subset(df2a, region!="Mexico")

#yourmodel <- lmer(y ~ treat + (1 | species) + (1 | author), data = df2) # this model has a singular fit
region.mod <- lmer(y ~ treat * region + (1 | study) , data = df2a) # this model runs fine at this point, but will produce singular fits when bootstrapped

summary(region.mod) # view summary output of model
AIC(region.mod) # check the AIC

```

**Bootstrap the model and estimate the 95% confidence intervals and mean**
```{r}
out <- simulate(region.mod,nsim=bootnum,seed=seed,re.form=NULL)
boots <- apply(out,2,function(x) predict(lmer(x ~ treat * region + (1 | study), data = df2a),re.form=NA))

boots <- cbind(predict(region.mod,re.form=NA), boots)
#estimate the mean and 95%confidence intervals for the prediction
df.reg <- (cbind(df2a, as.data.frame(t(apply(boots, 1, function(x) c(mean(x),quantile(x, c(0.025, 0.975))))))))

colnames(df.reg)[15:17] <- c("mean", "lowerci", "upperci")
head(df.reg)

```

###Plot estimated mean and CI
```{r echo=FALSE}
df.reg$treat <- factor(df.reg$treat, levels = c("yi.a","yi.w","yi.i"))

ggplot(df.reg, aes(x=treat, y=mean))+
  theme(panel.grid.major=element_blank(), legend.position="bottom", panel.grid.minor=element_blank(), panel.background=element_blank())+
  theme(axis.line=element_line(color="black"))+
  theme(legend.key=element_blank())+
  ylab("Mean effect size (LnRR)")+
  xlab("Treatment")+
  scale_x_discrete(labels=c("yi.a" = "acidification", "yi.w" = "warming", "yi.i" = "interaction"))+
  geom_hline(yintercept=0, linetype=2, colour="black")+
  geom_errorbar(aes(ymin=lowerci, ymax=upperci, width=0.1), size=0.6)+
  geom_point(size=6, shape=16, colour="black")+
  facet_grid(~region)+
  ylim(-1.6,0.6)
```

