---
title: "Results: Meta Analysis"
output:
  #pdf_document: default
  html_document: 
      theme: cerulean
      code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#htmltools::includeHTML("Bove_Results_MetaAnalysis.html")
```
*The following script and analyses were written by Colleen Bove and James Umbanhowar. This version of the script was last committed to GitHub on 08 January 2020*

In total, eleven studies met the standards of this meta-analysis, including the responses of thirteen Caribbean coral species collected from five different countries across the wider Caribbean (**Figure 1 and Supplementary Table S1**). Of the studies selected, only four performed fully factorial ocean acidification and warming experiments, and one performed two independent acidification and warming experiments. The most studied coral species from the Caribbean region were *Porites astreoides* (5 studies), *Acropora cervicornis* (4 studies), and *Siderastrea siderea* (4 studies). Finally, the Florida Keys and Belize were the two most-studied regions within the wider Caribbean fitting the criteria of this meta-analysis.

```{r packages and setup, include=FALSE}

library("metafor")
library("MAd")
library("ggplot2")
library("tidyr")
library("plyr")
library("lme4")
library("tidyverse")
library("tidystats")
library("RColorBrewer")
library("MuMIn")
library("Rmisc")
library("shiny")
library("plotly")
library("sjPlot")
library("sjmisc")
library("sjlabelled")
library("kableExtra")
library("cowplot")
#webshot::install_phantomjs()

date <- Sys.Date() #For saving with the crrent date
theme_set(theme_bw())
dodge<-position_dodge(0.2)
seed = 7 #seed to make results replicatable

```

```{r start, message=FALSE, warning=FALSE, echo = FALSE}

meta_tab <- read.csv("Table1_30Dec19.csv")
meta_tab$irradiance <- gsub("ND", NA, meta_tab$irradiance)
meta_tab$Feeding.1 <- gsub("ND", NA, meta_tab$Feeding.1)
meta_tab2 <- subset(meta_tab, Study != "Reneger et al. 2005")
meta_tab <- meta_tab2[, -c(3:13)]
meta_tab2 <- meta_tab2[, c(2:13, 20)]

meta_tab <- meta_tab %>% 
  rename(Feeding = Feeding.1, Seawater = Seawater.1)

df <- read.csv("effectsize_09May19.csv") 
df <- subset(df, author !="Jury_CALC")
df$units <- NULL
df$cite <- paste(df$author, df$year, sep = " ")
df <- subset(df, author != "Reneger")

calc <- read.csv("effectsize_arag_05Nov19.csv")  #for aragonite
calc <- subset(calc, units !="N")
calc$num <- 1:nrow(calc) # just helpful for the data manipulation
calc$units <- NULL
calc$author <- gsub("Jury_CALC", "Jury", calc$author)


calc<-calc %>% 
  left_join(meta_tab, by=c("author" = "author"))
calc <- calc[, -29]

df<-df %>% 
  left_join(meta_tab, by=c("author" = "author"))
df <- df[, -29]

```

####(a.)	Overall calcification response
Meta-analysis of the dataset revealed that calcification rates of Caribbean corals were negatively impacted by ocean acidification and warming independently (**Figure 2 and Supplementary Figures S1 – S6**). However, the 95% confidence interval of the combination of ocean warming and acidification overlapped zero, indicating no statistically clear trend towards synergistic or antagonistic effects of these treatments (**Figure 2 and Supplementary Table S2**). The resulting *Q~M~* indicates that treatments had different effects on calcification rates, however, the *Q~E~* suggests there is additional variation in the data outside the variance explained by the fixed effects in the analysis (**Supplementary Table S2**). 


<br/>

![**Figure 1** Coral collection sites of all included studies with experimental study represented by colour and treatments represented by shape: acidification only (circle), warming only (triangle), and the combination of acidification and warming (square). The lower left insert displays close-up of the Belize collection sites and the upper right insert displays the Florida Keys collection sites.](/Users/Colleen/Dropbox/Boston_Experiment/Manuscript3_metaanalysis/Code/Bove_MetaAnalysis/06Sep19 Update/FullMap_25April19.png)


<br/>

```{r table 1 of studies}

table1 <- kable(meta_tab2, booktabs = T, caption = "Table 1. Included Experimental Studies") %>%
  kable_styling(font_size = 10, full_width = F) %>% 
  column_spec(5, italic = T)
table1

#save_kable(table1, "Table1_Studies_26Sep19.pdf")

```
***

```{r ES1,  warning=FALSE, message=FALSE, tidy=TRUE}

# calculate ES and variance for ocean acidification
ES.oa<-escalc(measure="SMDH", m1i=cm1, m2i=am3, sd1i=s1, sd2i=s3, n1i=n1, n2i=n3, data=df)

# calculate ES and variance ES for ocean warming
ES.ow<-escalc(measure="SMDH", m1i=cm1, m2i=wm4, sd1i=s1, sd2i=s4, n1i=n1, n2i=n4, data=df)
ES.ow <- ES.ow %>% 
  rename(yi.w=yi, vi.w=vi)

#create DF of calculated variables of OA
acid<-data.frame("cm1"=ES.oa$cm1,"yi.a"=ES.oa$yi,"vi.a"=ES.oa$vi)

# merge together warming and acid DF 
ES.full<-merge(ES.ow,acid,by="cm1")
ES.full<-ES.full[order(ES.full$cite), ]
ES.full$yi.w<-(ES.full$yi.w)*-1
ES.full$yi.a<-(ES.full$yi.a)*-1

```

```{r sd}

# calculate sd for all 
ES.full$sd1<-ES.full$s1*sqrt(ES.full$n1)
ES.full$sd3<-ES.full$s3*sqrt(ES.full$n3)
ES.full$sd4<-ES.full$s4*sqrt(ES.full$n4)
ES.full$sd6<-ES.full$s6*sqrt(ES.full$n6)

# calculate S for all treatments
ES.full$Sa<-sqrt((ES.full$sd1^2/ES.full$n1)+(ES.full$sd3^2/ES.full$n3))
ES.full$Sw<-sqrt((ES.full$sd1^2/ES.full$n1)+(ES.full$sd4^2/ES.full$n4))
ES.full$Si<-sqrt((ES.full$sd1^2/ES.full$n1)+(ES.full$sd6^2/ES.full$n6))

```

```{r ESinter}

# calculate ES using Harvey et al equations for acidification, warming, and interaction studies
ES.full$lnRRa<-((ES.full$am3-ES.full$cm1)/ES.full$Sa)
ES.full$lnRRt<-((ES.full$wm4-ES.full$cm1)/ES.full$Sw)
ES.full$yi.i<-((ES.full$m6-ES.full$wm4)-(ES.full$am3-ES.full$cm1))/(2*ES.full$Si)

# calculate V for interaction studies ONLY
ES.full$vi.i<-(1/ES.full$n1)+(1/ES.full$n3)+(1/ES.full$n4)+(1/ES.full$n6)+((ES.full$yi.i^2)/(2*(ES.full$n1+ES.full$n3+ES.full$n4+ES.full$n6)))

```

```{r writeCSV}

csvFileName <- paste("ES.full", date, sep="_") 
csvFileName <- paste(csvFileName, ".csv", sep="")

write.csv(ES.full, file=csvFileName) 

```

```{r start2}

# read in the saved CSV and then removed any unneeded columns
df<-read.csv(csvFileName)
df<-df[,-c(2, 12:13, 16:18, 20:22, 24:26, 41:49)] #remove any unneeded columns

```

```{r datamanip}

#subset df for variance (v) values only as new df
df2b<-subset(df, select=c(X,vi.i,vi.a,vi.w))

#subset df to remove only variance (v) values as another new df
df.m <-subset(df, select=-c(vi.i,vi.a,vi.w))

# this creates df with treat and V as column headers in long format
df2b<-gather(df2b, treat, v,vi.i:vi.w)

#substitute y for v for merge later
df2b$treat<-gsub('v','y', df2b$treat)

#creates another long format df with treat and Y as headers (of Y values)
df.m <-gather(df.m, treat, y,yi.w:yi.i)
df.m $treat<-as.factor(df.m $treat)

#merge two df into one
df2<-merge(df.m, df2b, by=c("X", "treat"))

#remove any NAs
#df2<-na.omit(df2)
df2$study <- paste(df2$author, df2$year, sep = "_")
df2$X<-NULL

```

```{r}

df2$mag.p <- df2$a.pco2 - df2$c.pco2
df2$mag.t <- df2$w.temp - df2$c.temp

df2$mag.cent.p <- df2$mag.p - 127 # make a centered around 0 column
df2$mag.cent.t <- df2$mag.t - 1.0 # make a centered around 0 column

```

```{r rma,  message=FALSE, warning=FALSE}

# subset data by treatment
OW<-subset(df2, treat=="yi.w")
OA<-subset(df2, treat=="yi.a")
IN<-subset(df2, treat=="yi.i")

# fit model to each treatment
ow <- rma(y, v, data=OW)
oa <- rma(y, v, data=OA)
both <- rma(y, v, data=IN)

```


```{r wide to long, include=FALSE}

df2 <- gather(data = df2, key = "treat.mag", value = "mag", mag.p:mag.t)

```


```{r fullModel}

completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
}

df2 <- completeFun(df2, "y")
df2 <- completeFun(df2, "mag")

```

```{r global model selection,  message=FALSE, warning=FALSE}

g.mod.full <- rma.mv(y,v, mods = ~ treat * mag + days + irradiance + Feeding + Seawater + calc_method, random = list(~ (1|study), ~ (1|species)), data = df2) 
g.mod1 <- rma.mv(y,v, mods = ~ treat * mag + days + irradiance + Feeding + Seawater, random = list(~ (1|study), ~ (1|species)), data = df2) 
g.mod2 <- rma.mv(y,v, mods = ~ treat * mag + days + irradiance + Feeding, random = list(~ (1|study), ~ (1|species)), data = df2) # using this model
g.mod3 <- rma.mv(y,v, mods = ~ treat * mag + days + irradiance, random = list(~ (1|study), ~ (1|species)), data = df2) 
g.mod4 <- rma.mv(y,v, mods = ~ treat * mag, random = list(~ (1|study), ~ (1|species)), data = df2) 

g.mod5 <- rma.mv(y,v, mods = ~ treat * mag + irradiance + Feeding, random = list(~ (1|study), ~ (1|species)), data = df2) 
g.mod6 <- rma.mv(y,v, mods = ~ treat * mag + days + Feeding, random = list(~ (1|study), ~ (1|species)), data = df2) 
g.mod7 <- rma.mv(y,v, mods = ~ treat * mag + days + irradiance, random = list(~ (1|study), ~ (1|species)), data = df2) 

g.mod.d <- rma.mv(y,v, mods = ~ treat * mag + days, random = list(~ (1|study), ~ (1|species)), data = df2)
g.mod.i <- rma.mv(y,v, mods = ~ treat * mag + irradiance, random = list(~ (1|study), ~ (1|species)), data = df2)
g.mod.f <- rma.mv(y,v, mods = ~ treat * mag + Feeding, random = list(~ (1|study), ~ (1|species)), data = df2)
g.mod.s <- rma.mv(y,v, mods = ~ treat * mag + Seawater, random = list(~ (1|study), ~ (1|species)), data = df2)
g.mod.c <- rma.mv(y,v, mods = ~ treat * mag + calc_method, random = list(~ (1|study), ~ (1|species)), data = df2)

#AIC(g.mod.full,g.mod1 ,g.mod2 ,g.mod3 ,g.mod4, g.mod5, g.mod6, g.mod7,g.mod.d, g.mod.i, g.mod.f, g.mod.s, g.mod.c)

```

(model: mods = ~ treat * mag + days + irradiance + Feeding, random = list(~ (1|study), ~ (1|species))
```{r full model fit,  warning=FALSE}

global.mod <- rma.mv(y,v, mods = ~ treat * mag , random = list(~ (1|study), ~ (1|species)), data = df2) 

preds <- as.data.frame(predict(global.mod, newdata = df2))[,c(1,3:4)] # pull the mean and CI with the predict function
global.out <- cbind(df2, preds) # add columns to existing df for global output

global.mod

```

<br/><br/><br/>

```{r global mean estimation, include=FALSE}

pred <- summarySE(measurevar = "pred", groupvars = "treat", data = global.out)[,c(1,3)]
lb <- summarySE(measurevar = "ci.lb", groupvars = "treat", data = global.out)[,3]
up <- summarySE(measurevar = "ci.ub", groupvars = "treat", data = global.out)[,3]

pred.global <- cbind(pred, lb, up)

```


```{r plot full model,  fig.height = 3.5, fig.width = 4.5, warning=FALSE, fig.align="center"}

global.out$treat <- factor(global.out$treat, levels = c("yi.a","yi.w","yi.i"))
global.out$wt <- 1/(global.out$v)

global_CI <- 
  ggplot()+
  theme(panel.grid.major=element_blank(), legend.position="bottom", panel.grid.minor=element_blank(), panel.background=element_blank())+
  theme(axis.line=element_line(color="black"))+
  theme(legend.key=element_blank())+
  ylab("Mean effect size (SMD)")+
  xlab("")+
  #ggtitle("Figure 2. Global standardized mean difference by treatment")+
  scale_x_discrete(labels=c("yi.a" = "acidification", "yi.w" = "warming", "yi.i" = "combination"))+
  geom_hline(yintercept=0, linetype=2, colour="black")+
  geom_point(aes(x = treat, y = y, size = wt),  colour="grey", data = global.out, alpha = 0.5, position=position_jitter(width=0.1))+
  #geom_errorbar(aes(ymin=lb, ymax=ub, width=0.1), size=0.6)+
  geom_linerange(data = pred.global, aes(x = treat, ymin=lb, ymax=up), size=0.8)+
  #ylim(-1.6,0.6)+
  geom_point(data = pred.global, aes(x = treat, y = pred), size=10, shape=95)

div(ggplotly(global_CI), align = "center")

```

**Figure 2**
Mean effect (standard mean difference) and 95% confidence interval of ocean acidification, warming, and the combination of acidification and warming on calcification rate for all studies in the meta-analysis. Grey circles indicate the effect size of each individual study and the size of each circle represents the weight of each study (1/SE). Clear statistical evidence of a treatment effect is identified when the 95% confidence interval does not overlap zero. 


```{r save global output,  results = 'hide'}

pdf("Figure2_GlobalCI_30Dec19.pdf", width = 5, height = 3.5)
global_CI
dev.off()

```

```{r global model output table}

tidy.rma <- function(x, ...) {
  with(
    x,
    data.frame(
      estimate = b,
      std.error = se,
      z.value = zval,
      p.value = pval,
      conf.low = ci.lb,
      conf.high = ci.ub
    )
  )
}
tidy.rma.Q <- function(x, ...) {
  with(
    x,
    data.frame(
      QE = QE,
      p.value = QEp,
      QM = QM,
       p.value = QMp
    )
  )
}

global.Q <- tidy.rma.Q(global.mod)
global.Q <- data.frame(t(global.Q))
row.names(global.Q) <- c("QE", "p value", "QM", " p value")
colnames(global.Q) <- "estimate"
global.Q$std.error <- rep("", 4)
global.Q$z.value <- rep("", 4)
global.Q$p.value <- rep("", 4)
global.Q$conf.low <- rep("", 4)
global.Q$conf.high <- rep("", 4)


global.mod.tab <- tidy.rma(global.mod)
#row.names(global.mod.tab) <- c("acidification", "combination", "warming")
global.mod.tab$std.error <- round(global.mod.tab$std.error, 3)
global.mod.tab$z.value <- round(global.mod.tab$z.value, 3)
global.mod.tab$p.value <- round(global.mod.tab$p.value, 3)
global.mod.tab$conf.low <- round(global.mod.tab$conf.low, 3)
global.mod.tab$conf.high <- round(global.mod.tab$conf.high, 3)
global.mod.tab <- rbind(global.mod.tab, global.Q)

global.mod.tab <- global.mod.tab[,1:4]

table2 <- kable(global.mod.tab, digits = 2, caption = "Table 2. Gloabl Model Output") %>%
  kable_styling(font_size = 14, full_width = F) %>% 
  pack_rows("Model Results", 1, 6) %>%
  pack_rows("Variance Components", 7, 10)
table2

#save_kable(table2, "Table2_GlobalMod_26Sep19.pdf")

```


```{r parameter warming meta, include=FALSE}

w.global <- subset(df2, treat == "yi.w")
w.global <- subset(w.global, treat.mag == "mag.t")

w.global.mod <- rma.mv(y,v, mods = ~ mag + irradiance + Feeding + days, random = list(~ (1|study), ~ (1|species)), data = w.global) 

y <- summary(w.global.mod)$b
ci.lb <- summary(w.global.mod)$ci.lb
ci.ub <- summary(w.global.mod)$ci.ub
w.global.out <- data.frame(cbind(y,ci.lb,ci.ub))
colnames(w.global.out)[1]<-"Estimate"
colnames(w.global.out)[2]<-"Lower 95% CI"
colnames(w.global.out)[3]<-"Upper 95% CI"

rownames(w.global.out) <- c("Intercept", "Magnitude", "Irradiance", "Feeding", "Days")

```

```{r parameter acid meta, include=FALSE}

a.global <- subset(df2, treat == "yi.a")
a.global <- subset(a.global, treat.mag == "mag.p")
a.global.mod <- rma.mv(y,v, mods = ~ mag + irradiance + Feeding + days + mag:days, random = list(~ (1|study), ~ (1|species)), data = a.global) 

y <- summary(a.global.mod)$b
ci.lb <- summary(a.global.mod)$ci.lb
ci.ub <- summary(a.global.mod)$ci.ub
a.global.out <- data.frame(cbind(y,ci.lb,ci.ub))
colnames(a.global.out)[1]<-"Estimate"
colnames(a.global.out)[2]<-"Lower 95% CI"
colnames(a.global.out)[3]<-"Upper 95% CI"

rownames(a.global.out) <- c("Intercept", "Magnitude", "Irradiance", "Feeding","Days", "Magnitude:Days")

```

```{r parameter combination meta, include=FALSE}

i.global <- subset(df2, treat == "yi.i")
i.global.mod <- rma.mv(y,v, mods = ~ mag + irradiance + Feeding + days + mag:days, random = list(~ (1|study), ~ (1|species)), data = i.global) 

y <- summary(i.global.mod)$b
ci.lb <- summary(i.global.mod)$ci.lb
ci.ub <- summary(i.global.mod)$ci.ub
i.global.out <- data.frame(cbind(y,ci.lb,ci.ub))
colnames(i.global.out)[1]<-"Estimate"
colnames(i.global.out)[2]<-"Lower 95% CI"
colnames(i.global.out)[3]<-"Upper 95% CI"

rownames(i.global.out) <- c("Intercept", "Magnitude", "Irradiance", "Feeding","Days", "Magnitude:Days")

```

```{r parameter table}

full.param <- rbind(w.global.out, a.global.out, i.global.out)


param.tab <- kable(full.param, digits = 4, caption = "Table 1 | Effect Size of Experimental Design Parameters") %>% 
  kable_styling(font_size = 14, full_width = F) %>% 
  pack_rows("Warming Experiments", 1, 5) %>%
  pack_rows("Acidification Experiments", 6, 11) %>% 
  pack_rows("Combined Experiments", 12, 17)

param.tab

```


<br/>

####(b.)	Calcification response of Florida Keys versus Belize corals
Ocean acidification and warming independently reduced experimental calcification rates of corals from the Florida Keys. Furthermore, the combination of ocean acidification and warming resulted in a synergistic calcification rate reduction (**Figure 3A and Supplementary Table S3**). Conversely, corals from Belize only exhibited statistically clear reduced calcification rates under ocean warming (**Figure 3B and Supplementary Table 3**). The resulting *Q~M~* indicates that treatment and region both had different effects on coral calcification rates, however, *Q~E~* suggests there is additional variation in the data outside the variance explained by the fixed effects in the analysis (**Supplementary Tables S3**).

<br/>

```{r region model fit, warning=FALSE}

df2a<-subset(df2, region!="Curacao")
df2a<-subset(df2a, region!="Little Cayman")
df2a<-subset(df2a, region!="Mexico")

region.mod <- rma.mv(y,v, mods = ~ treat * region * mag, random = list(~ (1|study), ~ (1|species)), data = df2a)

preds2 <- as.data.frame(predict(region.mod, newdata = df2a))[,c(1,3:4)] # pull the mean and CI with the predict function
region.out <- cbind(df2a, preds2) # add columns to existing df for global output

region.mod

```


```{r region mean estimation, include=FALSE}

pred <- summarySE(measurevar = "pred", groupvars = c("treat", "region"), data = region.out)[,c(1,2,4)]
lb <- summarySE(measurevar = "ci.lb", groupvars = c("treat", "region"), data = region.out)[,4]
up <- summarySE(measurevar = "ci.ub", groupvars = c("treat", "region"), data = region.out)[,4]

pred.region <- cbind(pred, lb, up)

```


```{r plot region model,  warning=FALSE, fig.height = 3.7, fig.width = 7, fig.align="center"}

region.out$treat <- factor(region.out$treat, levels = c("yi.a","yi.w","yi.i"))
region.out$wt <- 1/(region.out$v)

region_CI <- ggplot()+
  theme(panel.grid.major=element_blank(), legend.position="bottom", panel.grid.minor=element_blank(), panel.background=element_blank())+
  theme(axis.line=element_line(color="black"))+
  theme(legend.key=element_blank())+
  ylab("Mean effect size (SMD)")+
  xlab("")+
  scale_x_discrete(labels=c("yi.a" = "acidification", "yi.w" = "warming", "yi.i" = "combination"))+
  geom_hline(yintercept=0, linetype=2, colour="black")+
  geom_point(aes(x = treat, y = y, size = wt),  colour="grey", data = region.out, alpha = 0.5, position=position_jitter(width=0.1))+
  #ggtitle("Figure 3.Standardized mean difference by treatment for Florida Keys and Southern Belize")+
  #geom_errorbar(aes(ymin=lb, ymax=ub, width=0.1), size=0.6)+
  geom_linerange(data = pred.region, aes(x = treat, ymin=lb, ymax=up), size=0.8)+
  geom_point(data = pred.region, aes(x = treat, y = pred), size=10, shape=95)+
  facet_wrap(~region)  

div(ggplotly(region_CI), align = "center")

```

**Figure 3** Mean effect (standard mean difference) and 95% confidence interval of ocean acidification, warming, and the combination of acidification and warming on calcification rate for (A) Florida Keys corals and (B) Belize corals. Grey circles indicate the effect size of each individual study and the size of each circle represents the weight of each study (1/SE). Clear statistical evidence of a treatment effect is identified when the 95% confidence interval does not overlap zero.

<br/>

```{r, regional model output table}

regional.Q <- tidy.rma.Q(region.mod)
regional.Q <- data.frame(t(regional.Q))
row.names(regional.Q) <- c("QE", "p value", "QM", " p value")
colnames(regional.Q) <- "estimate"
regional.Q$std.error <- rep("", 4)
regional.Q$z.value <- rep("", 4)
regional.Q$p.value <- rep("", 4)
regional.Q$conf.low <- rep("", 4)
regional.Q$conf.high <- rep("", 4)


region.mod.tab <- tidy.rma(region.mod)
#row.names(region.mod.tab) <- c("acidification", "combination", "warming","acidification ", "combination ", "warming ")
region.mod.tab$std.error <- round(region.mod.tab$std.error, 3)
region.mod.tab$z.value <- round(region.mod.tab$z.value, 3)
region.mod.tab$p.value <- round(region.mod.tab$p.value, 3)
region.mod.tab$conf.low <- round(region.mod.tab$conf.low, 3)
region.mod.tab$conf.high <- round(region.mod.tab$conf.high, 3)
region.mod.tab <- rbind(region.mod.tab, regional.Q)

region.mod.tab <- region.mod.tab[,1:4]

table3 <- kable(region.mod.tab, digits = 4, booktabs = T, caption = "Table 3. Regional Model Output") %>%
  kable_styling(font_size = 14, full_width = F) %>% 
  pack_rows("Model Results", 1, 12) %>%
  #pack_rows("Florida", 17, 20) %>%
  pack_rows("Variance Components", 13, 16)
table3

#save_kable(table3, "Table3_RegionalMod_26Sep19.pdf")
```

```{r save regional output,  results = 'hide'}

pdf("Figure3_RegionalCI_30Dec19.pdf", width = 8, height = 3.5)
region_CI
dev.off()

```


```{r}

calc.w<-calc[,-c(15:18, 23:27)] #remove columns not needed for warming analysis

# remove and NA columns from warming column
calc.w <- completeFun(calc.w, "wm4")

#subset calc for calcication rates
calc.b<-subset(calc.w, select=c(num, cm1, wm4))
# create df with treat and rate as column headers in long format
calc.b<-gather(calc.b, treat, rate, cm1:wm4)
# rename for useful treatment names
calc.b$treat<-gsub('cm1','control', calc.b$treat)
calc.b$treat<-gsub('wm4','warm', calc.b$treat)

#subset calc for SE
calc.c<-subset(calc.w, select=c(num, s1, s4))
# create df with treat and SE as column headers in long format
calc.c<-gather(calc.c, treat, SE, s1:s4)
# rename for useful treatment names
calc.c$treat<-gsub('s1','control', calc.c$treat)
calc.c$treat<-gsub('s4','warm', calc.c$treat)

#subset calc to remove only columns from above
calc.d <-subset(calc.w, select=-c(s1 ,s4 ,cm1, wm4, c.pco2, n1, n4))
# this creates df with treat and V as column headers in long format
calc.d<-gather(calc.d, treat, temp, c.temp:w.temp)
# rename for useful treatment names
calc.d$treat<-gsub('c.temp','control', calc.d$treat)
calc.d$treat<-gsub('w.temp','warm', calc.d$treat)

#merge dfs into one, final warming
calc.m<-merge(calc.d, calc.b, by=c("num", "treat"))
calc.warm<-merge(calc.m, calc.c, by=c("num", "treat"))

# create column for weight
calc.warm$wt <- 1 / calc.warm$SE

# create study column
calc.warm$study <- paste(calc.warm$author, calc.warm$year, sep = "_")

# center temps
calc.warm$cent.temp <- calc.warm$temp - mean(calc.warm$temp)
calc.warm$cent.pco2 <- rep("NA", 66)

```

 
```{r echo=TRUE, message=FALSE}

calc.warm$scale.temp <- scale(calc.warm$temp, center = TRUE, scale = TRUE)

warm.rate.region.mod <- lmer(rate ~ scale.temp + region + t.mag + (1|study) + (1|spec), weights = 1/SE, data = calc.warm,REML=F) #sing fit
warm.rate.region2.mod <- lmer(rate ~ scale.temp + region + I(scale.temp^2) + t.mag + (1|study) + (1|spec), weights = 1/SE, data = calc.warm,REML=F) 
warm.rate2.mod <- lmer(rate ~ scale.temp + I(scale.temp^2) + t.mag + (1|study) + (1|spec), weights = 1/SE, data = calc.warm,REML=F) 
warm.rate.mod <- lmer(rate ~ scale.temp + t.mag + (1|study) + (1|spec), weights = 1/SE, data = calc.warm,REML=F) 
warm.region.mod <- lmer(rate ~ region + t.mag + (1|study) + (1|spec), weights = 1/SE, data = calc.warm,REML=F) #sing fit

warm.rate2.day.mod <- lmer(rate ~ scale.temp + I(scale.temp^2) + t.mag + days + t.mag:days + (1|study) + (1|spec), weights = 1/SE, data = calc.warm,REML=F) 
warm.rate.day.mod <- lmer(rate ~ scale.temp + t.mag + days + t.mag:days + (1|study) + (1|spec), weights = 1/SE, data = calc.warm,REML=F) 

AICc(warm.rate.region.mod, warm.rate.region2.mod, warm.rate2.mod, warm.rate.mod, warm.region.mod,warm.rate2.day.mod,warm.rate.day.mod)


# best fit model without other predictors is warm.rate.region2.mod, with warm.rate2.day.mod as close second

summary(warm.rate.region2.mod) # view summary output of model
confint(warm.rate.region2.mod , method="boot")

warm.pred <- data.frame(temp=seq(26,32.01,length.out=100), region=rep(c("Belize", "Florida Keys", "Mexico"), length.out = 100), pred=predict(warm.rate.region2.mod, newdata = data.frame(scale.temp=seq(min(calc.warm$scale.temp),max(calc.warm$scale.temp),length.out=100), t.mag=seq(min(calc.warm$t.mag),max(calc.warm$t.mag),length.out=100), region=rep(c("Belize", "Florida Keys", "Mexico"), length.out = 100)), re.form=NA)) 

```

<br/>

####(c.)	Temperature impact on calcification rates across studies
Secondary analysis of mean calcification rates (mg cm^−2^ day^−1^) against treatment temperature across all studies revealed a parabolic response to temperature (**Figure 4 and Supplementary Tables S4, S5**). The parabolic trend was a result of treatment temperature rather than region, suggesting regional differences identified in the meta-analysis was a result of experimental temperatures selected.

<br/><br/><br/><br/><br/>


```{r  fig.align="center"}

global_OW_calc <-
ggplot(calc.warm, aes(x=temp, y=rate))+
  geom_point(aes(size=1/SE, shape=region, fill=region), colour="black")+
    theme(panel.grid.major=element_blank(), legend.position="right", panel.grid.minor=element_blank(), panel.background=element_blank())+
  theme(axis.line=element_line(color="black"))+
  theme(legend.key=element_blank())+
  scale_shape_manual("Region", values = c(21,22,24))+
  scale_fill_manual("Region", values = c("#5ab4ac", "#d8b365", "#f5f5f5"))+
  geom_line(data = warm.pred, aes(y = pred, group = region), size = 1)+
  ylab(bquote('Calcification rate (mg' ~cm^-2* ~day^-1*')'))+
  xlab(bquote('Temperature ('*degree*C*')'))
global_OW_calc

```

```{r save warming calc,  results = 'hide'}

pdf("Figure4_CalcWarm_30Dec19.pdf", width = 7, height = 4)
global_OW_calc
dev.off()

```

<br/>

**Figure 4.** Mean calcification rate (mg cm^−2^ day^−1^) of corals from each warming study by treatment temperature (°C) with the linear mixed effects model quadratic fit. Shape and colour of each point denotes study region (blue circle = Belize; brown square = Florida; grey triangle = Mexico) and size of shape represents the weight (1/SE) of study.


```{r lmer warming output,  fig.align='center' }

tab_model(warm.rate.region2.mod, show.r2 = FALSE, show.aicc = TRUE, title = "Table 4. Mixed-effects model output of effects of temperature on calcification rates")

```

<br/><br/>


```{r}

calc.a<-calc[,-c(19:27)] #remove columns not needed for acid analysis

calc.a$p.mag <- calc.a$c.pco2 - calc.a$a.pco2

completeFun <- function(data, desiredCols) {
  completeVec <- complete.cases(data[, desiredCols])
  return(data[completeVec, ])
} # remove and NA columns from acid column
calc.a <- completeFun(calc.a, "am3")

#subset calc for calcication rates
calc.b<-subset(calc.a, select=c(num, cm1, am3))
# create df with treat and rate as column headers in long format
calc.b<-gather(calc.b, treat, rate, cm1:am3)
# rename for useful treatment names
calc.b$treat<-gsub('cm1','control', calc.b$treat)
calc.b$treat<-gsub('am3','acid', calc.b$treat)

#subset calc for SE
calc.c<-subset(calc.a, select=c(num, s1, s3))
# create df with treat and SE as column headers in long format
calc.c<-gather(calc.c, treat, SE, s1:s3)
# rename for useful treatment names
calc.c$treat<-gsub('s1','control', calc.c$treat)
calc.c$treat<-gsub('s3','acid', calc.c$treat)

#subset calc to remove only columns from above
calc.d <-subset(calc.a, select=-c(s1 ,s3 ,cm1, am3, c.temp, n1, n3))
# this creates df with treat and V as column headers in long format
calc.d<-gather(calc.d, treat, acid, c.pco2:a.pco2)
# rename for useful treatment names
calc.d$treat<-gsub('c.pco2','control', calc.d$treat)
calc.d$treat<-gsub('a.pco2','acid', calc.d$treat)

#merge dfs into one, final acid
calc.m<-merge(calc.d, calc.b, by=c("num", "treat"))
calc.arag<-merge(calc.m, calc.c, by=c("num", "treat"))

# create column for weight
calc.arag$wt <- 1 / calc.arag$SE

# create study column
calc.arag$study <- paste(calc.arag$author, calc.arag$year, sep = "_")

# center temps
calc.arag$cent.pco2 <- calc.arag$acid - mean(calc.arag$acid)
#calc.arag$cent.pco2 <- rep("NA", 66)

```

```{r,  message=FALSE}

calc.arag$scale.arag <- scale(calc.arag$acid, center = TRUE, scale = TRUE)

acid.rate.region.mod <- lmer(rate ~ scale.arag + region + p.mag + (1|study) + (1|spec), weights = 1/SE, data = calc.arag,REML=F) 
acid.rate.region2.mod <- lmer(rate ~ scale.arag + region + p.mag + I(scale.arag^2) + (1|study) + (1|spec), weights = 1/SE, data = calc.arag,REML=F) 
#acid.rate2.mod <- lmer(rate ~ scale.arag + I(scale.arag^2) + p.mag + (1|study) + (1|spec), weights = 1/SE, data = calc.arag,REML=F) # didn't converge 
acid.rate.mod <- lmer(rate ~ scale.arag + p.mag + (1|study) + (1|spec), weights = 1/SE, data = calc.arag,REML=F) 
acid.region.mod <- lmer(rate ~ region + p.mag + (1|study) + (1|spec), weights = 1/SE, data = calc.arag,REML=F) 

acid.feed.day.mag.rate.mod <- lmer(rate ~ scale.arag + p.mag + Feeding + days + p.mag:days + (1|study) + (1|spec), weights = 1/SE, data = calc.arag,REML=F) 
acid.day.rate2.mod <- lmer(rate ~ scale.arag + I(scale.arag^2) + p.mag + days + p.mag:days + (1|study) + (1|spec), weights = 1/SE, data = calc.arag,REML=F) 
acid.feed.day.rate2.mod <- lmer(rate ~ scale.arag + I(scale.arag^2) + p.mag +  Feeding + days + (1|study) + (1|spec), weights = 1/SE, data = calc.arag,REML=F) 
acid.feed.rate2.mod <- lmer(rate ~ scale.arag + I(scale.arag^2) + p.mag +  Feeding + (1|study) + (1|spec), weights = 1/SE, data = calc.arag,REML=F) # best fit model


AICc(acid.rate.mod,acid.rate.region.mod,acid.rate.region2.mod,acid.region.mod,acid.feed.day.mag.rate.mod,acid.day.rate2.mod,acid.feed.day.rate2.mod,acid.feed.rate2.mod)

summary(acid.rate.region2.mod) # continuing with this one for simplicity
confint(acid.rate.region2.mod, method="boot")

acid.pred <- data.frame(acid=seq(0.7,5.8,length.out=100), region=rep(c("Belize", "Curacao", "Florida Keys"), length.out = 100),pred=predict(acid.rate.region2.mod, newdata = data.frame(scale.arag=seq(min(calc.arag$scale.arag),max(calc.arag$scale.arag),length.out=100), p.mag=seq(min(calc.arag$p.mag),max(calc.arag$p.mag), length.out=100), region=rep(c("Belize", "Curacao", "Florida Keys"), length.out=100)),re.form=NA))

```

  
```{r, arag region plot}

global_OA_calc <-
  ggplot(calc.arag, aes(x=acid, y=rate))+
  geom_point(aes(size=1/SE, shape=region, fill=region), colour="black")+
  theme(panel.grid.major=element_blank(), legend.position="right", panel.grid.minor=element_blank(), panel.background=element_blank())+
  theme(axis.line=element_line(color="black"))+
  theme(legend.key=element_blank())+
  scale_shape_manual("Region", values = c(21,22,24))+
  scale_fill_manual("Region", values = c("#5ab4ac", "#d8b365", "#f5f5f5"))+
  geom_line(data = acid.pred, aes(y = pred, group = region), size = 1)+
  ylab(bquote('Calcification rate (mg' ~cm^-2* ~day^-1*')'))+
  xlab(bquote('Aragonite Saturation State'))
global_OA_calc

```

```{r save acid calc,  results = 'hide'}

pdf("FigureS7_CalcAcid_06Nov19.pdf", width = 7, height = 4)
global_OA_calc
dev.off()

```


```{r lmer arag output,  fig.align='center' }
tab_model(acid.rate.region2.mod, show.r2 = FALSE, show.aicc = TRUE, title = "Table 4. Mixed-effects model output of effects of aragonite saturation state on calcification rates")
```

**Figure 5.** Mean calcification rate (mg cm−2 day−1) of corals from each acidification study by treatment aragonite saturation state (ΩArag) with the linear mixed effects model quadratic fit. Shape and colour of each point denotes study region (blue circle = Belize; brown square = Curacao; grey triangle = Florida Keys) and size of shape represents the weight (1/SE) of study.

***

### Supplemental Results
<br/>

#### I ) Ocean acidification forest and funnel plots
**S1.** Forest plot of ocean acidification study results.
```{r oa,  fig.height = 11, fig.width = 11, fig.align = "center"}

forest(oa, slab=paste(OA$cite, OA$spec, OA$site, sep=" - "), order=order(OA$cite))

```

**S2.** Funnel plot of ocean acidification study results.
```{r oa2,  fig.align = "center"}

funnel(oa, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend=TRUE)

```

```{r save oa forest,  results = 'hide'}

pdf("FigureS1_OAforest_26Sep19.pdf", width = 11, height = 12)
forest(oa, slab=paste(OA$cite, OA$spec, OA$site, sep=" - "), order=order(OA$cite))
dev.off()

pdf("FigureS2_OAfunnel_26Sep19.pdf", width = 7, height = 6)
funnel(oa, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend=TRUE)
dev.off()

```

#### II ) Ocean warming forest and funnel plots
**S3.** Forest plot of ocean warming study results.
```{r ow,  fig.height = 11, fig.width = 9, fig.align = "center"}

forest(ow, slab=paste(OW$cite, OW$spec, OW$site, sep=" - "), order=order(OW$cite))

```

**S4.** Funnel plot of ocean warming study results.
```{r ow2,  fig.height = 7, fig.width = 9, fig.align = "center"}

funnel(ow, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend=TRUE)

```

```{r save ow,  results = 'hide'}

pdf("FigureS3_OWforest_26Sep19.pdf", width = 9, height = 10)
forest(ow, slab=paste(OW$cite, OW$spec, OW$site, sep=" - "), order=order(OW$cite))
dev.off()

pdf("FigureS4_OWfunnel_26Sep19.pdf", width = 7, height = 5)
funnel(ow, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend="topleft")
dev.off()

```

#### III ) Combined ocean acidification and warming forest and funnel plots
**S5.** Forest plot of combination of ocean acidification and warming study results.
```{r both,  fig.height = 9, fig.width = 9, fig.align = "center"}

forest(both, slab=paste(IN$cite, IN$spec, IN$site, sep=" - "), order=order(IN$cite))


```

**S6.** Funnel plot of combination of ocean acidification and warming study results.
```{r both2,  fig.align = "center"}

funnel(both, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend=TRUE)

```

```{r save both,  results = 'hide'}

pdf("FigureS5_bothforest_26Sep19.pdf", width = 9, height = 8)
forest(both, slab=paste(IN$cite, IN$spec, IN$site, sep=" - "), order=order(IN$cite))
dev.off()

pdf("FigureS6_bothfunnel_26Sep19.pdf", width = 7, height = 6)
funnel(both, level=c(90, 95, 99), shade=c("white", "gray55", "gray75"), legend=TRUE)
dev.off()

```

```{r region1, include=FALSE}

df2a<-subset(df2, region!="Curacao")
df2a<-subset(df2a, region!="Little Cayman")
df2a<-subset(df2a, region!="Mexico")

# OA subset
OA2<-subset(df2a, treat=="yi.a")

# OW subset
OW2<-subset(df2a, treat=="yi.w")

# Interaction subset
IN2<-subset(df2a, treat=="yi.i")

```

```{r regionRMA, warning=FALSE, include=FALSE}

# acidification by warming
oa.r<-rma(y,v,mods=~region, data=OA2)

# warming by regions
ow.r<-rma(y,v,mods=~region, data=OW2)

# interactions by region
in.r<-rma(y,v,mods=~region, data=IN2)

```


```{r OAregion,  include=FALSE}

oa.r

```

```{r OWregion,  include=FALSE}

ow.r

```

```{r intregion,  include=FALSE}

in.r

```


```{r}
sessionInfo()
```

